What Happened to The Birth Rate?

### 2022-02-28 outputs/paper/paper.Rmd
backup progress

---
title: "What Happened to The Birth Rate?"
subtitle: "The Great Recession and the LGBT Movement"
author: 
  - Lantao Cui
  - Huakun Shen
  - Sijian Wu
thanks: "Replication avaliable at https://doi.org/10.48152/ssrp-mbxx-jm47. Code and data avaliable at https://github.com/cuilantao/Investigation-of-falling-US-Birth-Rates-since-the-great-recession"         
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper is based on Kearney's paper \"The Puzzle of Falling US Birth Rates since the Great Recession\". ^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, the US birth rate varied between a tight range of roughly 65 to 70 births per 1000 women between the age of 15 to 44. However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years. In this paper, we will expand our research on multiple countries and find out whether the level of impact from recession would be correlated with the changes in the birth rate for different countries. We will also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
header-includes:
  - \usepackage{subfig}
  - \usepackage{float}
  - \floatplacement{figure}{H}
---

@@

The percentage of LGBT (and many other gender identities) population has risen in recent years in the US \@ref(fig:fig-lgbt-trend-by-year), probably related to propaganda in social medias, lawmaking and political correctness.
This research doesn't try to investigate the causes of LGBT population growth, and since there is not much data for other genders, and LGBT are the majorities, we will focus on LGBT population and the potential effect they brought to birth rate.

"On February 12, 2004, the first same-sex marriage in the United States happened in San Francisco, California."
"On May 17, 2004, Massachusetts became the first state and sixth jurisdiction in the world to legalize same-sex marriage."
"In 2008, California and Connecticut both legalized same-sex marriage, followed by Iowa, Vermont, and New Hampshire. Up until 2012, legalization came through state courts, the enactment of state legislation, or the result of the decisions of federal courts. On November 6, 2012, Maine, Maryland and Washington became the first states to legalize same-sex marriage through popular vote."
"On June 26, 2015, in the landmark civil rights case Obergefell v. Hodges, the Supreme Court ruled that the fundamental right to marry is guaranteed to same-sex couples by both the Due Process Clause and the Equal Protection Clause of the Fourteenth Amendment to the United States Constitution."[@same_sex_marriage_states_2022]
The first two dates mentioned above represents the beginning of the legalization of same-sex marriage in the US, and is quite close to 2007, when the significant drop of birth rate began. There may be a composition of effects causing the significant drop of birth rate in 2007, and the legalization of same-sex population could potentially be one of main causes.

LGBT is the acronym of the following 4 gender identities, **L** for Lesbian, **G** for gay, **B** for bisexual, **T** for transgender. By definition, these genders should have a much lower birth rate than that of the heterosexual as normal births of human must depend on a male and a female. 
Figure \@ref(fig:fig-gender-raising-children-distribution) (b) shows the percentage of LGBT population that raise children. 71% of the LGBT population don't raise children. 
Then changes in birth rate and changes in LGBT population should have a negative proportional relationship (i.e. Growth in LGBT population leads to lower birth rate).
Thus, the population changes in these population are very likely to make an impact on the birth rate of the US.

@@

Figure \@ref(fig:fig-density-state) (a) and Figure \@ref(fig:fig-density-state) (b) shows the population density of Same Sex Couple and LGBT by States in the US [@will_institute].
Figure \@ref(fig:fig-density-state) [@cite_original] (c) shows the birth rate distribution by State in the US. Darker color means higher density or larger negative change in birth rate [@will_institute]. From these plots we can see that the states with higher LGBT density and same sex couples generally have larger negative change in birth rate. States on the west generally have darker color in all three plots. We cannot conclude whether LGBT population has a large or direct impact on US birth rate based on these plots yet, but there should be some correlation between birth rate and gender identity of Americans.

@@

With these information, we can conclude that the growth of LGBT population is able to cause a negative change in birth rate.

@@

Figure \@ref(fig:fig-LGBT-socioeconomic-indicators) shows some socioeconomic indicators classified by sexual orientation (LGBT or not), such as employment status, income level education level, and whether or not being insured and food insecure [@will_institute]. Darker color represents LGBT, lighter color represents non-LGBT.
From the plots, we can extract some socioeconomic indicators of LGBT. Comparing to non-LGBT population, LGBT population might be more likely to be unemployed, uninsured, being food insecure, having a low income (lower then $24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population). It's not rigorous to conclude for certain that LGBT are more likely to have these characteristics without a more in-depth statistical analysis, as there may be statistically insignificant errors in data. However, as all characteristics related to low-income discussed here are related to LGBT, there must be some correlation between these socioeconomic indicators and LGBT population and it's also worth noticing if government plans to affect birth rate.

@@

## LGBT and Cohort Effect

@@

>>>>

---
title: "What Happened to The Birth Rate?"
subtitle: "Impact to Birth Rate Brought By The Great Recession and The LGBT Movement"
author: 
  - Lantao Cui
  - Huakun Shen
  - Sijian Wu
thanks: "Replication avaliable at https://doi.org/10.48152/ssrp-mbxx-jm47. Code and data avaliable at https://github.com/cuilantao/Investigation-of-falling-US-Birth-Rates-since-the-great-recession"         
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "This paper is based on Kearney's paper \"The Puzzle of Falling US Birth Rates since the Great Recession\". ^[Original paper available at: https://www.aeaweb.org/articles?id=10.1257/app.2.2.95] [@cite_original] During the Great Recession, the US birth rate varied between a tight range of roughly 65 to 70 births per 1000 women between the age of 15 to 44. However, after the Great Recession, the birth rate started to decline significantly, it dropped by nearly 20 percent in 13 years. Maintaining the fertility rate to be around 2.1 (replacement level) is important because low fertility rate could lead to aging of population and economic stagnation of economy. In this paper, we expand our research on multiple countries and found out that not only the US was affected by the recession, but the whole World in the context of birth rate, and the most affected countries are mostly developing countries.
We also treat the rise in the LGBT community as a potential cause for the decline in birth rate and examine its effect, and found that the growth of LGBT population may have a cohort effect and lead caused a negative impact on the birth rate."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
header-includes:
  - \usepackage{subfig}
  - \usepackage{float}
  - \floatplacement{figure}{H}
---

@@

### Laws for Same Sex Marriage

"On February 12, 2004, the first same-sex marriage in the United States happened in San Francisco, California."
"On May 17, 2004, Massachusetts became the first state and sixth jurisdiction in the world to legalize same-sex marriage."
"In 2008, California and Connecticut both legalized same-sex marriage, followed by Iowa, Vermont, and New Hampshire. Up until 2012, legalization came through state courts, the enactment of state legislation, or the result of the decisions of federal courts. On November 6, 2012, Maine, Maryland and Washington became the first states to legalize same-sex marriage through popular vote."
"On June 26, 2015, in the landmark civil rights case Obergefell v. Hodges, the Supreme Court ruled that the fundamental right to marry is guaranteed to same-sex couples by both the Due Process Clause and the Equal Protection Clause of the Fourteenth Amendment to the United States Constitution."[@same_sex_marriage_states_2022].

The first two dates mentioned above represents the beginning of the legalization of same-sex marriage in the US, and is quite close to 2007, when the significant drop of birth rate began. There may be a composition of effects causing the significant drop of birth rate in 2007, and the legalization of same-sex population could potentially be one of main causes.

### LGBT Demographics

LGBT is the acronym of the following 4 gender identities, **L** for Lesbian, **G** for gay, **B** for bisexual, **T** for transgender. By definition, these genders should have a much lower birth rate than that of the heterosexual as normal births of human must depend on a male and a female. 
Then changes in birth rate and changes in LGBT population should have a negative proportional relationship (i.e. Growth in LGBT population leads to lower birth rate).
Thus, the population changes in these population are very likely to make an impact on the birth rate of the US.

@@

@@

### Socioeconomic Inicators

@@

Figure \@ref(fig:fig-LGBT-socioeconomic-indicators) shows some socioeconomic indicators classified by sexual orientation (LGBT or not), such as employment status, income level education level, and whether or not being insured and food insecure [@will_institute]. Darker color represents LGBT, lighter color represents non-LGBT.
From the plots, we can extract some socioeconomic indicators of LGBT. Comparing to non-LGBT population, LGBT population might be more likely to be unemployed, uninsured, being food insecure, having a low income (lower then $24k), and having a lower education level (percentage of LGBT population with Bachelor's or post-grad degree is slightly lower than that of the percentage of non-LGBT population).

### Geographical Distribution of LGBT Population and Birth Rate

Figure \@ref(fig:fig-density-state) (a) and Figure \@ref(fig:fig-density-state) (b) shows the population density of Same Sex Couple and LGBT by States in the US [@will_institute].
Figure \@ref(fig:fig-density-state) (c) [@cite_original] shows the birth rate distribution by state in the US. Darker color means higher density or larger negative change in birth rate [@will_institute]. From these plots we can see that the states with higher LGBT density and same sex couples generally have larger negative change in birth rate. States on the west generally have darker color in all three plots. We cannot conclude whether LGBT population has a large or direct impact on US birth rate based on these plots yet, but there should be some correlation between birth rate and gender identity of Americans.

@@

@@

## Cohort Effects and LGBT Movement Effects

With the results described in the **[Cohort Effect]** section, we can see that cohort effect can largely decrease the birth rate.

The percentage of LGBT (and many other gender identities) population has risen in recent years in the US \@ref(fig:fig-lgbt-trend-by-year), probably related to propaganda in social medias, lawmaking and political correctness.
This analysis doesn't try to investigate the causes of LGBT population growth, and since there is not much data for other genders, and LGBT are the majorities, we will focus on LGBT population and the potential effect they brought to birth rate.

With the results described in the **[Effect of LGBT Movement]** section, we showed that many factors related to LGBT population growth is able to cause a negative change in birth rate.

In section **[Socioeconomic Inicators]** we also discussed some socioeconomic characteristics of the LGBT population. The characteristics are all related to low-income or low-education level. However, it is not rigorous to conclude for certain that LGBT are more likely to have these characteristics without a more in-depth statistical analysis, as there may be statistically insignificant errors in data. However, as all characteristics related to low-income discussed here are related to LGBT, there must be some correlation between these socioeconomic indicators and LGBT population and it's also worth noticing if government wishes to adjust birth rate.
